---
title: "Build a simple DAG"
author: "Malte Braitmaier"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build a simple DAG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(simple.dag)
```

This vignette shows how to build a DAG using the base R plotting commands. Example data and helper functions are provided with the simple.dag package. 

## Load example data 

Nodes are specified in a dara.frame. Check for instance this example dataset: 

```{r}
data(nodes) 
nodes
```

Edges between nodes are specified in a separate data.frame. Additional graphics parameters can be specified in this data.frame (currently lwd and col): 

```{r}
data(edges) 
edges 
```

## Pre-processing of data 

In order for the DAG to look somewhat nice, the edges should not start directly on the nodes, but there should be a little buffer around them. For this, the define.arrow function can be used: 

```{r}
data(edges) 
add.to.edges <- vector(mode = "list", length = nrow(edges))
for (i in 1:length(add.to.edges)) {
  add.to.edges[[i]] <- define.arrow(
    x.0 = nodes[nodes$name == edges$from[i],"x"], 
    y.0 = nodes[nodes$name == edges$from[i],"y"], 
    x.1 = nodes[nodes$name == edges$to[i],"x"], 
    y.1 = nodes[nodes$name == edges$to[i],"y"], 
    radius = 0.2
  )
}
add.to.edges <- do.call(rbind, add.to.edges)
edges <- cbind(edges, add.to.edges)
edges 
```

## Use basic plotting functions to build the DAG 

Now use base R to plot the DAG: 

```{r}
par(mar = c(.1, .1, .1, .1))
plot(c(-1,-1), c(1,1), ylim = c(-1.1, 1.1), xlim = c(-1.1, 1.1), 
     axes = F, xlab = "", ylab = "", type = "n")
text(nodes$x, nodes$y, parse(text = nodes$description))
for (i in 1:nrow(edges)) {
  arrows(
    edges$x.0[i], 
    edges$y.0[i], 
    edges$x.1[i], 
    edges$y.1[i], 
    length = .1, 
    lwd = edges$lwd[i], 
    col = edges$col[i]
  )
}
```
